---
title: "Dirchlet Multinomial Simulation"
author:
  - name: Kenong Su
    email: kenong.su@pennmedicine.upenn.edu
shorttitle: DM simulation
abstract: >
 This includes basic steps for simulating DM distribution in spatial transcriptomics data. 
vignette: >
  %\VignetteIndexEntry{The DM Simulation Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\usepackage[utf8]{inputenc}
output:
  BiocStyle::html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    fig_width: 5
---
<style type="text/css">
  pre:not([class]) {
    background-color: white;
  }
  .tocify-subheader > .tocify-item {
  text-indent: 1px;}
</style>


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

\vspace{.1in}

# Read the data
Using Method of Moments (MoM) approach to estimate $N$ and $P$ parameters in multinomial distributions. 

```{r read data, message=FALSE, results='hide'}
library(SingleCellExperiment)
library(zellkonverter)
library(MCMCpack)
IDC = readH5AD("../data/IDC.h5ad")
regions = IDC$region
Y = as.matrix(assays(IDC)[[1]])
```

# Estimation
Input count matrix and estimate parameters

```{r estimation}
print(IDC)
source("Estimation_DM.R")
est_alpha = estPara(Y, regions)
```

# Simulate the new library size
The sequence depth is $50\%$ shrinked by using gamma distribution. Specifically, we model the shrinkage effect by using gamma(25, 50), where the mean is 0.5, and sd is 0.1. 

```{r seq dep}
nsample = ncol(Y)
seqDepth = colSums(Y)
set.seed(123)
newseqDepth = seqDepth*rgamma(nsample, 25, 50)
newseqDepth = round(newseqDepth)
```

# Simulation
```{r simulation}
Ynew = matrix(0, ncol = nsample, nrow = nrow(Y))
for (i in 1:nsample){
    reg = regions[i]
    tmp_alpha = est_alpha[, reg]
    PP = rdirichlet(1, alpha = tmp_alpha)
    tmp_SeqDep = newseqDepth[i]
    XX = rmultinom(1, size = tmp_SeqDep, prob = PP)
    Ynew[, i] = XX
}
IDC2 = SingleCellExperiment(assays = list(counts = Ynew), colData  = colData(IDC), rowData = rowData(IDC))
print(IDC)
writeH5AD(IDC2, file = "IDC_down.h5ad")
```

# Simulation (simple)
```{r simulation2}
set.seed(123)
ngene = nrow(Y)
shrinkages = matrix(rgamma(nsample*ngene, 25, 50), nrow = ngene, ncol=nsample)
Ynew = Y * shrinkages
IDC3 = SingleCellExperiment(assays = list(counts = Ynew), colData  = colData(IDC), rowData = rowData(IDC))
writeH5AD(IDC3, file = "IDC_down_2.h5ad")
```